Estimating Correspondences of Deformable Objects ``In-the-wild"

### 2015-11-03 contents/introduction.tex

\item The abundance of complex visual data, spread mostly through the internet via web services such as Youtube, Flickr and Google Images. The latter has led to the development of large databases of "in-the-wild" images of human faces and bodies \cite{Belhumeur2011, Le2012, Zhu2012, Burgos2013}. Equally important is the task of manual annotation of images with regards to face and body parts and that has been undertaken by several research teams \cite{sagonas_iccv_300w_2013}.

@@

@@

There are two drawbacks on building deformable models that rely on landmarks:
\begin{itemize}

\item Annotating with regards to consistent landmark annotation is an extremely time consuming, tedious and labour intensive work \cite{sagonas_iccv_300w_2013}, which is usually performed by a trained person. Furthermore, for many objects it requires a highly skilled person to identify and annotate a set of images with regards to a consistent set of landmarks. For example human ears have very complicated inner structures (e.g. helix, crus antihelicis, scapha, tragus, lobe etc.) which differs remarkably between ears. Moreover, certain ear parts, such as  Fossa Triangularies and Crus Helicis, do not appear in all ears and  their visibility is highly sensitive to pose and illumination variation.

\item The nature of many deformable objects does not allow them to be consistently annotated with regards to a consistent set of landmarks (e.g. bottles, fruits etc.). Also the boundary/outline of objects, such as faces, ears, body can be very difficultly annotated consistently, Figure \ref{fig:intro}. That is why many state-of-the-art methods opt to leave the boundary out for comparison \cite{Tzimiropoulos2014, Asthana2014}. The majority of the state-of-the-art methods for model-based landmark localisation \cite{Cao2012, Zhu2012, Xiong2013, Tzimiropoulos2014, Asthana2014} are not applicable for objects with inconsistent sets of landmarks.

@@

In this paper, we provide a solution for annotating an object with regards to its deformations that requires considerable less effort and in the same time can be used to define statistical deformations for objects without consistent set of landmarks. That is, we argue that it is better to annotate an object with regards to a set of continuous lines that describe the shape of the object. An example is provided in Figure \ref{fig:intro} where the shape of the facial parts is annotated with continuous lines instead of a number of discrete landmarks. Similarly, the shape annotations for the ear, a bottle and human body is shown in Figure \ref{fig:intro} Annotating facial images by following this procedure takes 42, 38 and 43 secs per image for faces, ears and bottles correspondingly. Furthermore, these shapes can be automatically produced by recent methods that perform discriminative segmentation of objects.

@@

We show that this way dense correspondences between images can be established and dense Active Appearance Models (AAMs) can be build on top of that. We show that the proposed methodology even though not tailored to landmark localization can be applied for landmark localization achieving very good performance. Finally, we demonstrate the performance of the proposed methodology for describing objects that do not have consistent landmarks. To the best of our knowledge this is the first methodology which can create a dense statistical model for AAM which can operate in "in-the-wild".

@@

Summarizing, the contributions proposed in this paper are:
\begin{itemize}
  \item Constructing SDMs using training data with inconsistent annotation.
  \item Introduce highly effective curve annotation which tremendously reduces manual work overload but maintains same performance level against state-of-the-art SDMs.
  \item Building dense correspondent SDMs with shape flow
\end{itemize}

>>>>

\item The abundance of complex visual data, spread mostly through the Internet via web services such as Youtube, Flickr and Google Images. The latter has led to the development of large databases of "in-the-wild" images of human faces and bodies \cite{Belhumeur2011, Le2012, Zhu2012, Burgos2013}.

@@

\item Equally important is the task of manual annotation of images with regards to face and body parts and that has been undertaken by several research teams \cite{sagonas_iccv_300w_2013,charles2013domain,dantone2014body,andriluka14cvpr}.

@@

There are two drawbacks on building deformable models that directly rely on landmarks:
\begin{itemize}

\item Annotating with regards to consistent landmark annotation is an extremely time consuming, tedious and labour intensive work \cite{sagonas_iccv_300w_2013}, which is usually performed by a trained person . Furthermore, for many objects it requires a highly skilled person to identify and annotate a set of images with regards to a consistent set of landmarks. For example human ears have very complicated inner structures (e.g. helix, crus antihelicis, scapha, tragus, lobe etc.) which differs remarkably between ears. Moreover, certain ear parts, such as  Fossa Triangularies and Crus Helicis, do not appear in all ears and  their visibility is highly sensitive to pose and illumination variation. Another example is human body which is generally annotated with regards to a number of landmarks that intuitively correspond to a set of body joints. The task of annotation was undertaken by a crowd-sourcing Internet marketplace, so-called Amazon Mechanical Turk (AMT). Unfortunately, these annotations are, in many cases, not consistent\footnote{From personal experience the quality of annotations that have been produced from AMT in case of faces are extremely inaccurate and cannot by any means compared with the ones provided by the recent competition \cite{sagonas_iccv_300w_2013}} (please see Figure \ref{fig:intro}). As it was also pointed out recently, the inconsistencies in body joint annotations may also render the comparison between different human pose estimation methodologies irrelevant \cite{tompson2015efficient}.

\item The nature of many deformable objects does not allow them to be consistently annotated with regards to a consistent set of landmarks (e.g. bottles, fruits etc.). Also the boundary/outline of objects, such as faces, ears, body can be very difficultly annotated consistently. That is why many state-of-the-art methods opt to leave the boundary out for comparison \cite{Tzimiropoulos2014, Asthana2014}. The majority of the state-of-the-art methods for model-based landmark localisation \cite{Cao2012, Zhu2012, Xiong2013, Tzimiropoulos2014, Asthana2014} are not applicable for objects with inconsistent sets of landmarks.

@@

In this paper, we provide a solution for annotating an object with regards to its deformations that requires considerable less effort and in the same time can be used to define statistical deformations for objects without consistent set of landmarks. We apply the proposed method to construct deformable models of the outline of  human body parts (e.g., arms and legs). To the best of our knowledge this is the first SDM of the outline of human body parts. The proposed SDM can be also used to provide accurate and consistent annotations of several of the body joints (such as wrist etc.).

To this end, we argue and empirically demonstrate that it is better to annotate an object with regards to a set of continuous lines that describe the shape of the object. An example is provided in Figure \ref{fig:intro} where the shape of the human arm is annotated with continuous lines instead of a number of discrete landmarks. Similarly, the shape annotations for the ear, a bottle and human body is shown in Figure \ref{fig:intro} Annotating facial images by following this procedure takes 42, 38 and 43 secs per image for arm, faces, ears and bottles correspondingly. Furthermore, these shapes can be automatically produced by recent methods that perform discriminative segmentation of objects.

@@

We show that this way dense correspondences between images can be established and dense Active Appearance Models (AAMs) can be build on top of that. We show that the proposed methodology even though not tailored to landmark localization can be applied for landmark localization achieving very good performance. We demonstrate the performance of the proposed methodology for describing objects that do not have consistent landmarks. Furthermore, We show that by sampling the shape of the dense AAM a part-based AAM can be trained with its parts being in correspondence. We show that such a part-based AAM can used to produce state-of-the-art results in localisation of human body parts in challenging database, outperforming state-of-the-art methods [..]. Finally, we show that the proposed part-based AAM can be used to provide consistent annotations for certain body parts.

%To the best of our knowledge this is the first methodology which can %create a dense statistical model for AAM which can operate in %"in-the-wild".

@@

Summarizing, the contributions proposed in this paper are:
\begin{itemize}

  \item We propose one of the first methodologies that construct SDMs from a set of training data with inconsistent annotation. We show that the proposed methodology, since it uses highly effective curve annotation, tremendously reduces manual work overload but maintains same performance level against state-of-the-art SDMs. We show that by sampling the dense shape of SDM very powerful part-based SDMs can

  \item We propose the first SDMs of the outline of human parts, such as human arms, which can be applied for both highly accurate body part segmentation and pose estimation, as well as for providing accurate annotations for body joints.
